---
title: "Monte Carlo Simulation of a Hypothesis Test Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Monte Carlo Simulation of a Hypothesis Test Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
notice: "@*"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  warning  = FALSE,
  message  = FALSE,
  fig.width  = 7,
  fig.height = 6,
  fig.align  = "center"
)
```

```{r setup}
library(Survive-380-Anna-the-best)
```

This vignette demonstrates how to estimate the Type I error rate, Type II error rate, and power of a hypothesis test for binomially distributed samples using Monte Carlo simulation, via three functions from the `Survive-380-Anna-the-best` package:

- Estimate Type I error `Survive-380-Anna-the-best::mc_type1_binom`

- Estimate Type II error and power when alternative is well-define `Survive-380-Anna-the-best::mc_power_simple_binom`

- Estimate Type II error and power curve when alternative is not well-define `Survive-380-Anna-the-best::mc_power_curve_binom`
