---
title: "Monte Carlo Simulation of a Hypothesis Test Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Monte Carlo Simulation of a Hypothesis Test Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
nocite: "@*"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(fig.width=7, fig.height=6,fig.align  = "center") 
```

```{r setup}
library(Survive380AnnaTheBest)
```

This vignette demonstrates how to estimate the Type I error rate, Type II error rate, and power of a hypothesis test for binomially distributed samples using Monte Carlo simulation, via four functions from the `Survive380AnnaTheBest` package:

- Calculates the p_values and standard error using Monte Carlo sampling `Survive380AnnaTheBest::mc_pval_binom`

- Estimate Type I error `Survive380AnnaTheBest::mc_type1_binom`

- Estimate Type II error and power at a specific $p_{1}$ `Survive380AnnaTheBest::mc_power_simple_binom`

- Estimate power curve across a grid of $p$-value `Survive380AnnaTheBest::mc_power_curve_binom`

We will produce:

1. A summary table of Type I, Type II error and power

2. A histogram of simulated $p$-value with a vertical line at $\alpha$

3. A power curve plot

## Quick assumption check

Since we are using normal approximation to perform z-test, a common assumption is both $np_{0}$ and $n(1-p_{0})$ should be reasonably big(often $\ge 10$)

```{r}
p_0 <- 0.5
n <- 50

c(np0 = n * p_0, n1p0 = n * (1 - p_0))
```

## Estimated p-value

### Computing p_val

```{r}
set.seed(2000)
p_0 <- 0.5 # success probability
sample_size <- 15 # number of trials
alternative <- "less"
s_obs <- 10 # s_obs: number of success observed in one data set

p_value <- mc_pval_binom(
  p_0 = p_0,
  s_obs = s_obs,
  sample_size = sample_size,
  alternative = alternative
)

p_value

# once the function has been changed: 
# we now have a vector of p-values from simulation
# change the variable name to "p_values" to match code in ggplot below

```

### Histogram

```{r badplot}
# DELETE THIS R CODE BLOCK ONCE WE HAVE FIXED THE FUNCTION
# THE GOOD PLOT IS IN THE BLOCK BELOW

library(ggplot2)

# we want to use p-value to make a histogram of the simulated p-values
# adjust the variable names once the function code has been adjusted

# we have the p-value from the function, calculated above, add in the p-value as a line.  
# need to simulate the graph using dbinom, number of successes vs probability
# dbinom should have a seed set, which we defined earlier

x_vals <- 0:sample_size  # All possible values (0 successes to n successes)
probs <- dbinom(x_vals, size = sample_size, prob = p_0) # theoretical distribution

# Create data frame for plotting
plot_data <- data.frame(x = x_vals, px= probs)

# the following plot is incorrect, the p-value line does not work since p-value is not an x-value; the line on the plot should instead be the significance threshold
ggplot(plot_data, aes(x, px)) + 
  geom_col(fill ="cornflowerblue") +
  xlab("x, number of successes") +
  ylab("p(x)") + ylim(0, 1) +
  labs(title = "Histogram of Binomial Distribution",
       subtitle = paste("number of trials =", sample_size, ", probability of success=", p_0, ", \n", "number of success observed =", s_obs, ", alternative =", alternative)) + 
  geom_vline(aes(xintercept = p_value$p_value), color = "coral", linewidth=1)+ 
  annotate(x = p_value$p_value, y = +Inf, label = paste("p-value:", round(p_value$p_value, digits = 4)), vjust = 2, geom = "label")

# this plot works for one-tailed less; what about greater and two-sided?
# add an if statement
```

```{r goodplot}
# HISTOGRAM OF SIMULATED P-VALUES WITH SIGNIFICANCE LEVEL ALPHA

######
# 1. simulate the p-values in a vector form so ggplot can be run 
# take this out once function code is working 
  N = 10^5

  # 1. standardize s_obs
  p_hat <- s_obs / sample_size
  z_obs <- (p_hat - p_0) / sqrt((p_0 * (1 - p_0)) / sample_size)
  
  # 2. simulate data under H_0
  x <- rbinom(N, size = sample_size, prob = p_0)
  
  # 3. compute z-statistic under normal approximation
  p_hat <- x / sample_size
  z <- (p_hat - p_0)/sqrt((p_0 * (1 - p_0)) / sample_size)
  
  # 4. Estimating p_value using Monte Carlo sampling
  p_values <- pnorm(abs(z)) # if alternative is less
  # we get a vector of pvalues

######
# remove the above once the function code is working
  
library(ggplot2)
  
alpha <- 0.05  # significance level
# should we put alpha as a parameter in the function? since i think we said that was something that could be controlled by the user. 

rejections <- (p_values < alpha) # for checking rejection region (FYI)

# NEED AN IF STATEMENT FOR DIFFERENT ALTERNATIVES
ggplot(data.frame(p_value = p_values), aes(x = p_value)) +
  geom_histogram(bins = 30, fill = "cornflowerblue", color = "black", alpha = 0.7) +
  geom_vline(xintercept = alpha, color = "red", linetype = "dashed", linewidth = 1) +
  annotate(x = alpha, y = +Inf, label = paste("α =", alpha), vjust = 2, geom = "label", size = 3, color = "red") +
  # Add shaded rejection region
  annotate("rect", xmin = 0, xmax = alpha, ymin = 0, ymax = Inf, alpha = 0.2, fill = "darkgoldenrod2") +
  annotate(x = alpha/2, y = +Inf, label = paste("Rejection", "\n", "Region"), vjust = 3, geom = "label", size = 2.5, color = "darkgoldenrod2") +
  labs(title = "Histogram of Simulated P-Values",
       subtitle = paste("Monte Carlo Simulation of One-Sample Z-test for Proportions", "\n" ,"of Binomially Distributed Samples"),
       caption = paste("number of trials =", sample_size, ", probability of success=", p_0, ", \n", "number of success observed =", s_obs, ", alternative =", alternative, ", \n", "proportion of p-values < α:", mean(p_values < alpha)),
       x = "P-value",
       y = "Frequency") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

```


## Estimated Type I error

```{r}
set.seed(1)
p_0 <- 0.5
sample_size <- 50
alternative <- "greater"

type1_rate <- mc_type1_binom(
  p_0 = p_0,
  sample_size = sample_size,
  alternative = alternative
)

type1_rate
```


## Estimated Type II error and power for a specific alternative value $p_{1}$

```{r}
set.seed(1)
p_0 <- 0.9
p_1 <- 0.1
sample_size <- 60
alternative <- "less"

power_type2 <- mc_power_simple_binom(
  p_0 = p_0,
  p_1 = p_1,
  sample_size = sample_size,
  alternative = alternative
)

power_type2
```

### Power Curve when alternative is not well-defined

```{r}
set.seed(1)
p_0 <- 0.6
sample_size <- 70
alternative <- "two_sided"

power_curve <- mc_power_curve_binom(
  p_0 = p_0,
  sample_size = sample_size,
  alternative = alternative
)

power_curve
```

### draw the plot of power curve 
```{r}
library(ggplot2)
ggplot(power_curve, aes(x = p_true, y = power_hat)) +
  geom_line(color = "dodgerblue", linewidth = 1.2) +
  geom_point(color = "dodgerblue", size = 2) +
  geom_errorbar(aes(ymin = power_hat - 2*power_mc_se, 
                    ymax = power_hat + 2*power_mc_se), 
                width = 0.01, color = "grey50", alpha = 0.5) +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "red") +
  annotate("text", x = min(power_curve$p_true), y = 0.08, 
           label = "Significance Level (0.05)", color = "red", hjust = 0, size = 3) +
  labs(title = "Power Curve of Binomial Test",
       subtitle = paste("Numbers of trails =", sample_size, ", Probability of success =", p_0, ",Alternative =", alternative),
       x = "True Probability (p)",
       y = "Power (Probability of Rejecting H0)") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))
```


## Bibliography

## Generative AI Usage Statement

Generative AI was used to debugging part of Function_Code, 