---
title: "Monte Carlo Simulation of a Hypothesis Test Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Monte Carlo Simulation of a Hypothesis Test Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
nocite: "@*"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(fig.width=7, fig.height=6,fig.align  = "center") 
```

```{r setup}
library(Survive380AnnaTheBest)
```

This vignette demonstrates how to estimate the Type I error rate, Type II error rate, and power of a hypothesis test for binomially distributed samples using Monte Carlo simulation, via four functions from the `Survive380AnnaTheBest` package:

- Calculates the p_values and standard error using Monte Carlo sampling `Survive380AnnaTheBest::mc_pval_binom`

- Estimate Type I error `Survive380AnnaTheBest::mc_type1_binom`

- Estimate Type II error and power at a specific $p_{1}$ `Survive380AnnaTheBest::mc_power_simple_binom`

- Estimate power curve across a grid of $p$-value `Survive380AnnaTheBest::mc_power_curve_binom`

We will produce:

1. A summary table of Type I, Type II error and power

2. A histogram of simulated $p$-value with a vertical line at $\alpha$

3. A power curve plot

## Quick assumption check

Since we are using normal approximation to perform z-test, a common assumption is both $np_{0}$ and $n(1-p_{0})$ should be reasonably big(often $\ge 10$)

```{r}
p_0 <- 0.5
n <- 50

c(np0 = n * p_0, n1p0 = n * (1 - p_0))
```

## Estimated p-value

### Computing p_val

```{r}
set.seed(2000)
p_0 <- 0.5 # success probability
sample_size <- 15 # number of trials
alternative <- "less"
s_obs <- 10 # s_obs: number of success observed in one data set

p_value <- mc_pval_binom(
  p_0 = p_0,
  s_obs = s_obs,
  sample_size = sample_size,
  alternative = alternative
)

p_value
```


### Histogram of p value

```{r}
# please add up the code to show a histogram of simulated p-values, with a vertical line indicating the chosen significance level $\alpha$ (might using ggplot2 to do so)

library(ggplot2)

# we want to use p-value to make a histogram of the simulated p-values
# adjust the variable names once the function code has been adjusted

# we have the p-value from the function, calculated above, add in the p-value as a line.  
# need to simulate the graph using dbinom, number of successes vs probability
# dbinom should have a seed set, which we defined earlier

x_vals <- 0:sample_size  # All possible values (0 successes to n successes)
probs <- dbinom(x_vals, size = sample_size, prob = p_0) # theoretical distribution

# Create data frame for plotting
plot_data <- data.frame(x = x_vals, px= probs)


ggplot(plot_data, aes(x, px)) + 
  geom_col(fill ="cornflowerblue") +
  xlab("x, number of successes") +
  ylab("p(x)") + ylim(0, 1) +
  labs(title = "Histogram of Binomial Distribution",
       subtitle = paste("number of trials =", sample_size, ", probability of success=", p_0, ", \n", "number of success observed =", s_obs, ", alternative =", alternative)) + 
  geom_vline(aes(xintercept = p_value$p_value), color = "coral", linewidth=1)+ 
  annotate(x = p_value$p_value, y = +Inf, label = paste("p-value:", round(p_value$p_value, digits = 4)), vjust = 2, geom = "label")

# this plot works for one-tailed less; what about greater and two-sided?
# add an if statement

```


## Estimated Type I error

```{r}
set.seed(1)
p_0 <- 0.5
sample_size <- 50
alternative <- "greater"

type1_rate <- mc_type1_binom(
  p_0 = p_0,
  sample_size = sample_size,
  alternative = alternative
)

type1_rate
```


## Estimated Type II error and power for a specific alternative value $p_{1}$

```{r}
set.seed(1)
p_0 <- 0.9
p_1 <- 0.1
sample_size <- 60
alternative <- "less"

power_type2 <- mc_power_simple_binom(
  p_0 = p_0,
  p_1 = p_1,
  sample_size = sample_size,
  alternative = alternative
)

power_type2
```

### Power Curve when alternative is not well-defined

```{r}
set.seed(1)
p_0 <- 0.6
sample_size <- 70
alternative <- "two_sided"

power_curve <- mc_power_curve_binom(
  p_0 = p_0,
  sample_size = sample_size,
  alternative = alternative
)

power_curve
```

### draw the plot of power curve  (two sided)
```{r}
library(ggplot2)

ggplot(power_curve, aes(x = p_true, 
                        y = power_hat)) +
  geom_line(color = "dodgerblue", 
            linewidth = 1.2) +
  geom_errorbar(aes(ymin = power_hat - 2*power_mc_se, 
                    ymax = power_hat + 2*power_mc_se), 
                width = 0.01, 
                color = "grey50", 
                alpha = 0.5) +
  geom_hline(yintercept = 0.05, 
             linetype = "dashed", color = "red") +
  annotate("text", x = min(power_curve$p_true), 
           y = 0.08, label = "Significance Level (0.05)", 
           color = "red", 
           hjust = 0, 
           size = 3) +
  labs(title = "Power Curve of Binomial Test",
       subtitle = paste("Numbers of trails =", sample_size, 
                        ", Probability of success =", p_0, 
                        ",Alternative =", alternative),
       x = "True Probability (p)", 
       y = "Power (Probability of Rejecting H0)") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))
```

### Power Curve when alternative is greater

```{r}
set.seed(1)
p_0 <- 0.6
sample_size <- 70

alternative_one_sided <- "greater" 

power_curve_one_sided <- mc_power_curve_binom(
  p_0 = p_0,
  sample_size = sample_size,
  alternative = alternative_one_sided
)

power_curve_one_sided
```

### draw the plot of power curve  (greater)

```{r}
library(ggplot2)

colnames(power_curve_one_sided) <- c("p_true", "power")
alternative <- alternative_one_sided 

ggplot(data = power_curve_one_sided, aes(x = p_true, y = power)) +
  geom_line(color = "dodgerblue", linewidth = 1.2) +
  geom_vline(xintercept = p_0, linetype = "dashed", color = "gray50", linewidth = 1) + 
  annotate("text", x = p_0 + 0.02, y = 0.1, label = paste("p0 =", p_0), color = "gray50", angle = 90) +
  labs(
    title = "One-Tailed Power Curve for Binomial Test",
    subtitle = paste("Number of trials =", sample_size, 
                     ", Null Proportion =", p_0, 
                     ", Alternative =", alternative),
    x = "True Underlying Proportion (p)",
    y = "Statistical Power"
  ) +
  theme_minimal()
```

### Power vs. Sample Size Curve

```{r}
#This indicates the sample size required to achieve the desired test capability. Here,   we use 80% as an example.

#set initials
sample_sizes <- seq(10, 200, by = 10)
M <- 1000
p_true <- 0.65  
p_null <- 0.5
alpha <- 0.05
power_results <- numeric(length(sample_sizes))

for(i in 1:length(sample_sizes)) {
  n <- sample_sizes[i]
  p_hat <- rbinom(M, n, p_true) / n
  z_stats <- (p_hat - p_null) / sqrt(p_null * (1 - p_null) / n)
  p_values <- 2 * pnorm(-abs(z_stats))
  power_results[i] <- mean(p_values < alpha)
}
df_power <- matrix(c(sample_sizes, power_results),
                   ncol = 2,
                   dimnames = list(NULL, c("SampleSize", "Power")))
achieved_target <- df_power[df_power[, "Power"] >= 0.8, , drop = FALSE]
if (nrow(achieved_target) > 0) {
  required_n <- achieved_target[1, "SampleSize"]
  cat("To achieve 80% test power，we need at least n =", required_n, "\n")
} else {
  cat("fail to reach 80% test power，please enhance the sample size \n")
}
```

### Plot of Power vs. Sample Size Curve

```{r}
library(ggplot2)
ggplot(as.data.frame(df_power), aes(x = SampleSize, y = Power)) +
  geom_line(color = "dodgerblue", linewidth = 1.2) +
  
  geom_hline(yintercept = 0.8, 
             linetype = "dashed", 
             color = "red", 
             linewidth = 1.2) +
  
  annotate("text", x = min(df_power[, "SampleSize"]), 
           y = 0.82, 
           label = "Target Power: 80%", color = "red", hjust = 0) +
  
  labs(title = " Sample Size vs. Power",
       x = "Sample Size (n)",
       y = "Statistical Power") +
  scale_y_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
  theme_minimal()
```
## Bibliography

## Generative AI Usage Statement

Generative AI was used to debugging part of Function_Code, 