---
title: "Monte Carlo Simulation of a Hypothesis Test Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Monte Carlo Simulation of a Hypothesis Test Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
nocite: "@*"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(fig.width=7, fig.height=6,fig.align  = "center") 
```

```{r setup}
library(Survive380AnnaTheBest)
```

This vignette demonstrates how to estimate the Type I error rate, Type II error rate, and power of a hypothesis test for binomially distributed samples using Monte Carlo simulation, via four functions from the `Survive380AnnaTheBest` package:

- Calculates the p_values and standard error using Monte Carlo sampling `Survive380AnnaTheBest::mc_pval_binom`

- Estimate Type I error `Survive380AnnaTheBest::mc_type1_binom`

- Estimate Type II error and power at a specific $p_{1}$ `Survive380AnnaTheBest::mc_power_simple_binom`

- Estimate power curve across a grid of $p$-value `Survive380AnnaTheBest::mc_power_curve_binom`

We will produce:

1. A summary table of Type I, Type II error and power

2. A histogram of simulated $p$-value with a vertical line at $\alpha$

3. A power curve plot

## Quick assumption check

Since we are using normal approximation to perform z-test, a common assumption is both $np_{0}$ and $n(1-p_{0})$ should be reasonably big(often $\ge 10$)

```{r}
p_0 <- 0.5
n <- 50

c(np0 = n * p_0, n1p0 = n * (1 - p_0))
```

## Estimated p-value

### Computing p_val under a obervation

```{r}
set.seed(1)
p_0 <- 0.5 # success probability
sample_size <- 50 # number of trials
alternative <- "less"
s_obs <- 20 # s_obs: number of success observed in one data set

p_value <- mc_pval_binom(
  p_0 = p_0,
  s_obs = s_obs,
  sample_size = sample_size,
  alternative = alternative
)

p_value
```

### Histogram

```{r}
# following code used to compute null p_value under H_0
p_0 <- 0.3
sample_size <- 60
alternative <- "two_sided"

p_values <- mc_pval_null(
  p_0 = p_0,
  sample_size = sample_size,
  alternative = alternative
)

p_values


library(ggplot2)
  
alpha <- 0.05  # significance level
# should we put alpha as a parameter in the function? since i think we said that was something that could be controlled by the user. 

rejections <- (p_values < alpha) # for checking rejection region (FYI)

# NEED AN IF STATEMENT FOR DIFFERENT ALTERNATIVES
ggplot(data.frame(p_value = p_values), aes(x = p_value)) +
  geom_histogram(bins = 30, fill = "cornflowerblue", color = "black", alpha = 0.7) +
  geom_vline(xintercept = alpha, color = "red", linetype = "dashed", linewidth = 1) +
  annotate(x = alpha, y = +Inf, label = paste("α =", alpha), vjust = 2, geom = "label", size = 3, color = "red") +
  # Add shaded rejection region
  annotate("rect", xmin = 0, xmax = alpha, ymin = 0, ymax = Inf, alpha = 0.2, fill = "darkgoldenrod2") +
  annotate(x = alpha/2, y = +Inf, label = paste("Rejection", "\n", "Region"), vjust = 3, geom = "label", size = 2.5, color = "darkgoldenrod2") +
  labs(title = "Histogram of Simulated P-Values",
       subtitle = paste("Monte Carlo Simulation of One-Sample Z-test for Proportions", "\n" ,"of Binomially Distributed Samples"),
       caption = paste("number of trials =", sample_size, ", probability of success=", p_0, ", \n", "number of success observed =", s_obs, ", alternative =", alternative, ", \n", "proportion of p-values < α:", mean(p_values < alpha)),
       x = "P-value",
       y = "Frequency") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

```


## Estimated Type I error

```{r}
set.seed(1)
p_0 <- 0.5
sample_size <- 50
alternative <- "greater"

type1_rate <- mc_type1_binom(
  p_0 = p_0,
  sample_size = sample_size,
  alternative = alternative
)

type1_rate
```


## Estimated Type II error and power for a specific alternative value $p_{1}$

```{r}
set.seed(1)
p_0 <- 0.9
p_1 <- 0.1
sample_size <- 60
alternative <- "less"

power_type2 <- mc_power_simple_binom(
  p_0 = p_0,
  p_1 = p_1,
  sample_size = sample_size,
  alternative = alternative
)

power_type2
```

### Power Curve when alternative is not well-defined

```{r}
set.seed(1)
p_0 <- 0.6
sample_size <- 70
alternative <- "two_sided"

power_curve <- mc_power_curve_binom(
  p_0 = p_0,
  sample_size = sample_size,
  alternative = alternative
)

power_curve
```

### draw the plot of power curve  (two sided)
```{r}
library(ggplot2)

ggplot(power_curve, aes(x = p_true, 
                        y = power_hat)) +
  geom_line(color = "cornflowerblue", 
            linewidth = 1.2) +
  geom_errorbar(aes(ymin = power_hat - 2*power_mc_se, 
                    ymax = power_hat + 2*power_mc_se), 
                width = 0.01, 
                color = "grey50", 
                alpha = 0.5) +
  geom_hline(yintercept = 0.05, 
             linetype = "dashed", color = "red") +
  annotate("text", x = min(power_curve$p_true), 
           y = 0.08, label = "Significance Level (0.05)", 
           color = "red", 
           hjust = 0, 
           size = 3) +
  labs(title = "Power Curve of Binomial Test",
       subtitle = paste("Numbers of trails =", sample_size, 
                        ", Probability of success =", p_0, 
                        ",Alternative =", alternative),
       x = "True Probability (p)", 
       y = "Power (Probability of Rejecting H0)") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))
```

### Power Curve when alternative is greater

```{r}
set.seed(1)
p_0 <- 0.6
sample_size <- 70

alternative_one_sided <- "greater" 

power_curve_one_sided <- mc_power_curve_binom(
  p_0 = p_0,
  sample_size = sample_size,
  alternative = alternative_one_sided
)

power_curve_one_sided
```

### draw the plot of power curve  (greater)

```{r}
library(ggplot2)

colnames(power_curve_one_sided) <- c("p_true", "power")
alternative <- alternative_one_sided 
alpha_level <- 0.05

ggplot(data = power_curve_one_sided, aes(x = p_true, y = power)) +
  geom_line(color = "cornflowerblue", linewidth = 1.2) +
  geom_vline(xintercept = p_0, linetype = "dashed", color = "coral", linewidth = 1) +    annotate("text", x = p_0 + 0.02, y = alpha_level+0.05, 
           label = paste("p0 =", p_0), color = "coral", angle = 90) +
  geom_hline(yintercept = alpha_level, linetype = "dashed", color = "red") +
  annotate("text", x = p_0 + 0.02, y = alpha_level + 0.05, 
           label = paste("Alpha =", alpha_level), color = "red", size = 3) +
  labs(
    title = "One-Tailed Power Curve for Binomial Test",
    subtitle = paste("Number of trials =", sample_size, 
                     ", Probability of success =", p_0, 
                     ", Alternative =", alternative),
    x = "True Underlying Proportion (p)",
    y = "Statistical Power"
  ) +
  theme_minimal()
```

### Power vs. Sample Size Curve

```{r}
#This indicates the sample size required to achieve the desired test capability. Here,   we use 80% as an example.

#set initials
sample_sizes <- seq(10, 200, by = 10)
M <- 1000
p_true <- 0.65  
p_null <- 0.5
alpha <- 0.05
power_results <- numeric(length(sample_sizes))

for(i in 1:length(sample_sizes)) {
  n <- sample_sizes[i]
  p_hat <- rbinom(M, n, p_true) / n
  z_stats <- (p_hat - p_null) / sqrt(p_null * (1 - p_null) / n)
  p_values <- 2 * pnorm(-abs(z_stats))
  power_results[i] <- mean(p_values < alpha)
}
df_power <- matrix(c(sample_sizes, power_results),
                   ncol = 2,
                   dimnames = list(NULL, c("SampleSize", "Power")))
achieved_target <- df_power[df_power[, "Power"] >= 0.8, , drop = FALSE]
if (nrow(achieved_target) > 0) {
  required_n <- achieved_target[1, "SampleSize"]
  cat("To achieve 80% test power，we need at least n =", required_n, "\n")
} else {
  cat("fail to reach 80% test power，please enhance the sample size \n")
}
```

### Plot of Power vs. Sample Size Curve

```{r}
library(ggplot2)
ggplot(as.data.frame(df_power), aes(x = SampleSize, y = Power)) +
  geom_line(color = "cornflowerblue", linewidth = 1.2) +
  
  geom_hline(yintercept = 0.8, 
             linetype = "dashed", 
             color = "red", 
             linewidth = 0.8) +
  
  annotate("text", x = min(df_power[, "SampleSize"]), 
           y = 0.82, 
           label = "Target Power: 80%", color = "red", hjust = 0) +
  
  labs(title = " Sample Size vs. Power",
       x = "Sample Size (n)",
       y = "Statistical Power") +
  scale_y_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
  theme_minimal()
```

## Bibliography

## Generative AI Usage Statement

Generative AI was used to debug part of Function_Code, and assist in plot creation using ggplot2. 