---
title: "Monte Carlo Simulation of a Hypothesis Test Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Monte Carlo Simulation of a Hypothesis Test Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
nocite: "@*"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(fig.width=7, fig.height=6,fig.align  = "center") 
```

```{r setup}
library(Survive380AnnaTheBest)
```

This vignette demonstrates how to estimate the Type I error rate, Type II error rate, and power of a hypothesis test for binomially distributed samples using Monte Carlo simulation, via four functions from the `Survive380AnnaTheBest` package:

- Calculates the p_values and standard error using Monte Carlo sampling `Survive380AnnaTheBest::mc_pval_binom`

- Estimate Type I error `Survive380AnnaTheBest::mc_type1_binom`

- Estimate Type II error and power at a specific $p_{1}$ `Survive380AnnaTheBest::mc_power_simple_binom`

- Estimate power curve across a grid of $p$-value `Survive380AnnaTheBest::mc_power_curve_binom`

We will produce:

1. A summary table of Type I, Type II error and power

2. A histogram of simulated $p$-value with a vertical line at $\alpha$

3. A power curve plot

## Quick assumption check

Since we are using normal approximation to perform z-test, a common assumption is both $np_{0}$ and $n(1-p_{0})$ should be reasonably big(often $\ge 10$)

```{r}
p_0 <- 0.5
n <- 50

c(np0 = n * p_0, n1p0 = n * (1 - p_0))
```

## Estimated p-value

### Computing p_val

```{r}
set.seed(2000)
p_0 <- 0.5 # success probability
sample_size <- 15 # number of trials
alternative <- "less"
s_obs <- 10 # s_obs: number of success observed in one data set

p_value <- mc_pval_binom(
  p_0 = p_0,
  s_obs = s_obs,
  sample_size = sample_size,
  alternative = alternative
)

p_value
```

### Histogram

```{r}
# please add up the code to show a histogram of simulated p-values, with a vertical line indicating the chosen significance level $\alpha$ (might using ggplot2 to do so)

library(ggplot2)

# we want to use p-value to make a histogram of the simulated p-values
# adjust the variable names once the function code has been adjusted

# we have the p-value from the function, calculated above, add in the p-value as a line.  
# need to simulate the graph using dbinom, number of successes vs probability
# dbinom should have a seed set, which we defined earlier

x_vals <- 0:sample_size  # All possible values (0 successes to n successes)
probs <- dbinom(x_vals, size = sample_size, prob = p_0) # theoretical distribution

# Create data frame for plotting
plot_data <- data.frame(x = x_vals, px= probs)


ggplot(plot_data, aes(x, px)) + 
  geom_col(fill ="cornflowerblue") +
  xlab("x, number of successes") +
  ylab("p(x)") + ylim(0, 1) +
  labs(title = "Histogram of Binomial Distribution",
       subtitle = paste("number of trials =", sample_size, ", probability of success=", p_0, ", \n", "number of success observed =", s_obs, ", alternative =", alternative)) + 
  geom_vline(aes(xintercept = p_value$p_value), color = "coral", linewidth=1)+ 
  annotate(x = p_value$p_value, y = +Inf, label = paste("p-value:", round(p_value$p_value, digits = 4)), vjust = 2, geom = "label")

# this plot works for one-tailed less; what about greater and two-sided?
# add an if statement

```

## Estimated Type I error

```{r}
set.seed(1)
p_0 <- 0.5
sample_size <- 50
alternative <- "greater"

type1_rate <- mc_type1_binom(
  p_0 = p_0,
  sample_size = sample_size,
  alternative = alternative
)

type1_rate
```


## Estimated Type II error and power for a specific alternative value $p_{1}$

```{r}
set.seed(1)
p_0 <- 0.9
p_1 <- 0.1
sample_size <- 60
alternative <- "less"

power_type2 <- mc_power_simple_binom(
  p_0 = p_0,
  p_1 = p_1,
  sample_size = sample_size,
  alternative = alternative
)

power_type2
```

## Power Curve when alternative is not well-defined

```{r}
set.seed(1)
p_0 <- 0.6
sample_size <- 70
alternative <- "two_sided"

power_curve <- mc_power_curve_binom(
  p_0 = p_0,
  sample_size = sample_size,
  alternative = alternative
)

power_curve

# also draw the power curve here
```


## Bibliography

## Generative AI Usage Statement

Generative AI was used to debugging part of Function_Code, 